\documentclass[orivec]{llncs}

\title{A class of automata for the verification of infinite, resource-allocating behaviours%
 \thanks{Research partially funded by projects EU 
 ASCENS (nr. 257414), EU QUANTICOL (nr. 600708), 
 IT MIUR CINA and PAR FAS 2007-2013 Regione Toscana TRACE-IT.}
}
\author{Vincenzo Ciancia\inst{1} \and Matteo Sammartino\inst{2}}
\institute{ISTI-CNR, Pisa \and Dipartimento di Informatica, Universit\`a di Pisa, Pisa }
\input{macros}
\input{matteo-macros}
\input{vincenzo-macros}

\raggedbottom
\frenchspacing

\begin{document}

\maketitle

\begin{abstract}
Process calculi for service-oriented computing often feature generation of fresh resources. So-called nominal automata have been studied both as semantic models for such calculi, and as acceptors of languages of finite words over infinite alphabets. In this paper we investigate nominal automata that accept infinite words. These automata are a generalisation of deterministic Muller automata to the setting of nominal sets. We prove decidability of complement, union, intersection, emptiness and equivalence, and determinacy by ultimately periodic words. The key to obtain such results is to use finite representations of the (otherwise infinite-state) defined class of automata. The definition of such operations enables model checking of process calculi featuring infinite behaviours, and resource allocation, to be implemented using classical automata-theoretic methods.
% 
%  Nominal sets, presheaf categories, and named sets have successfully served as models of the state space of process calculi featuring resource generation. More recently, automata built in such categories have been studied as acceptors of languages of finite words over infinite alphabets. In this paper we investigate automata whose state spaces are nominal sets, and that accept infinite words. These automata are a generalisation of deterministic Muller automata to the setting of nominal sets. 
%  %Remarkably, the obtained languages are determined by their ultimately periodic fragments, as in the classical case. 
%  We prove decidability of complement, union, intersection, emptiness and equivalence. This is shown using finite representations of the (otherwise infinite-state) defined class of automata. The definition of such operations enables model checking of systems featuring infinite behaviours, and resource allocation, to be implemented using automata-theoretic methods, like in the classical case.
\end{abstract}

\section{Introduction}\label{sec:introduction}\input{introduction1}

\section{Example: peer-to-peer system}\label{sec:example}
\input{example}

\section{Background}\label{sec:background}\input{background}

\section{Nominal \texorpdfstring{$\omega$}{omega}-regular languages}\label{sec:languages}\input{languages}


\section{Finite automata}\label{sec:hd-automata}
\input{hd-automata}


\section{Synchronized product and boolean operations}\label{sec:sync-product}
\input{synchronized-product}
%
%\section{Boolean operations and decidability}\label{sec:boolean-operations-decidability}
\input{boolean-operations-decidability}

\section{Ultimately-periodic words}\label{sec:up-words}\input{up-words}

\section{Conclusions}\label{sec:conclusions}
\input{future-work}
\input{related-work}

\paragraph{Acknowledgements.} The authors thank Nikos Tzevelekos, Emilio Tuosto and Gianluca Mezzetti for several fruitful discussions related to nominal automata.

\bibliographystyle{splncs}
\bibliography{bibliography}
%\input{biblio}

\appendix
\section{Proofs}
\input{proofs}

\end{document}
